---
title: "p8105_hw1_yj2581"
author: "YucongJiang"
date: "2019-9-16"
output: github_document
---

```{r setup, include = FALSE}
# load the necessary package
library(tidyverse)
```

## Problem 1

#### Create the data frame `problem_1` based on the instruction

```{r dataframe_create_problem1}
set.seed(2581) # set the randomized seed to ensure the reproducibility

problem_1 <- tibble(
  vec_numeric = rnorm(8),
  vec_logical = vec_numeric > 0,
  vec_character = c("Tom", "Sam", "Bob", "Ivy", "Hans", "Ann", "Tony", "Roy"),
  vec_grade = factor(c("A", "B", "C", "B", "B", "A", "C", "B"))
)
```

#### Take the mean of each variable in the dataframe

  * The mean of `vec_numeric` is `r mean(pull(problem_1, vec_numeric))`.
  * The mean of `vec_logical` is `r mean(pull(problem_1, vec_logical))`.
  * The mean of `vec_character` is `r mean(pull(problem_1, vec_character))`.
  * The mean of `vec_grade` is `r mean(pull(problem_1, vec_grade))`.
  
We can see that:

  * It works when taking the mean of `vec_numeric` and `vec_logical`.
  * It doesn't work when taking the mean of `vec_character` and `vec_grade`.

#### Apply `as.numeric` function to the logical, character and factor variables

```{r type_transform, eval = FALSE}
as.numeric(pull(problem_1, vec_logical))
as.numeric(pull(problem_1, vec_character))
as.numeric(pull(problem_1, vec_grade))
```

We can see that:

  * For `vec_logical`, the values `FALSE` and `TRUE` are transformed to `0` and `1` respectively.
  * For `vec_character`, it doesn't work and the results are all `NA`.
  * For `vec_grade`, it returns the underlying integer representation. Reference from R Documentation.
  
These results explain what happens when try to take the mean:

  * For `vec_logical`, it is coerced first and then taken the mean.
  * For `vec_character`, it doesn't work since the character cannot be converted to the numeric.
  * For `vec_grade`, the integer may not correspond to the levels, so the function refuses to take the mean.
  
#### Convert the logical vector to different types and multiply the random sample by the results

```{r convert and multiply, warning = FALSE}
# Show the results and hide the warning message
logical_to_numeric <- as.numeric(pull(problem_1, vec_logical)) *
                      pull(problem_1, vec_numeric)
logical_to_numeric

logical_to_factor <- as.factor(pull(problem_1, vec_logical)) *
                     pull(problem_1, vec_numeric)
logical_to_factor

logical_factor_numeric <- as.numeric(as.factor(pull(problem_1, vec_logical))) *
                          pull(problem_1, vec_numeric)
logical_factor_numeric
```

For `logical_to_factor`, it warns '*' not meaningful for factors.

## Problem 2

#### Create the dataframe based on the instruction

```{r dataframe_create_problem2}
set.seed(8125) # set the randomized seed to ensure the reproducibility

problem_2 <- tibble(
  x = rnorm(500),
  y = rnorm(500),
  logical_indicator = x + y > 1,
  numeric_indicator = as.numeric(logical_indicator),
  factor_indicator = as.factor(numeric_indicator)
)
```

#### Write a short description of your vector using inline R code

  * The dataset has `r nrow(problem_2)` rows and `r ncol(problem_2)` columns.
  * The mean of `x` is `r mean(pull(problem_2, x))`.
  * The median of `x` is `r median(pull(problem_2, x))`.
  * The standard deviation of `x` is `r sd(pull(problem_2, x))`.
  * The proportion of cases for which `x + y > 1` is `r sum(pull(problem_2, numeric_indicator))/500`.

#### Make 3 scatterplots of `y` vs `x` based on the instruction; save the first scatterplot

```{r scatterplot}
ggplot(problem_2, aes(x = x, y = y, color = logical_indicator)) + geom_point()
ggsave("scatterplot_1.png") # save the first scatterplot

ggplot(problem_2, aes(x = x, y = y, color = numeric_indicator)) + geom_point()

ggplot(problem_2, aes(x = x, y = y, color = factor_indicator)) + geom_point()
```

  * For the first and third scatterplots, the variables are discrete, so the plots use 2 colors: 
  red represents `FALSE` or `0` and cyan represents `TRUE` or `1`.
  * For the second scatterplot, the variables are continuous, so the color scales from black to blue 
  to represent the variables from 0 to 1